\bf On Devon Allen's Disqualification at the 2022 World Track and Field Championships

### 2024-11-20 Code/RankComparison.Rmd

clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_hurdle_2022,
                method = "ds",
                exact = TRUE, B = 1000000)
```

@@

#clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_2023,
 #               method = "ds",
  #              exact = TRUE, B = 1000000)
```

>>>>

clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_2022,
                method = "ds",
                exact = TRUE, B = 1000000)
```

@@

clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_2023,
                method = "ds",
                exact = TRUE, B = 1000000)
```

### 2024-11-20 Manuscript/reply.tex

@@

Thank you for these suggestions, we were able to expand the scope of our data
by including results from the men's and women's 100 meter dash.  We considered
adding additional sprinting events such as the 200 meter dash but our analysis
suggested that reaction times were slower in the 200 meter dash.  After running
a model that included a coefficient for event type on a data set that combined
the men's 110 meter hurdle and 100 meter dash, the event term was not significant.
However after we re-ran the model on the 110 meter hurdle, 100 and 200 meter dash,
the effect became significant. When we peformed a TukeyHSD test to compare the
times in the three categories: 100 meter dash, 110 meter hurdles, 200 meter dash,
we found a p-value of 0.0020625 comparing the means of the 200 meter dash and
the 110 meter hurdlesand a p-value of 0.0000174 comparing the means of the 100
meter dash and the 200 meter dash.  This indicated to us that the reaction times
for the 200 meter were too slow to fit into our study on extreme reaction times
and thus we chose to exclude 200 meter data.

This is not particularly suprising as athletes who compete in the longer
events may not have as good of a start as those in the 100
meter events because reaction time plays a much smaller role in the outcome of
the race. While it is common for amateur and collegiate athletes to compete in
multiple sprinting events, it seems reasonable that a world championship 100
meter runner may be only a very good 200 meter runner.  We are trying to
estimate the probabilty of observing a extreme reaction time to determine
whether 0.1 seconds is fair but if we include data from competitions with slower
reaction times, then 0.1 seconds is going to be more extreme.

\jy{At the end of each reply, we need to state what changes we made to
  the manuscript with their specific locations. The explanation here
  (the exclusion of 200m dash) should be included in the paper as well.}

@@

The heat effect measures the speed-level of the heat.  We added some additional
explanations in the first paragraph of section 3.2 to make it clear what the
venue and heat effect refer to. We did not want to call it a "race effect" as we
were worried that may be interpreted to mean the race of the athlete.  We did
not think it would be informative to  add and effect for individual runner's
average ability as many of observations in the data come from runners
with very few (1-3) data points, so taking an average of such a small sample may
not be informative, especially if there are outliers or an athlete had a bad
start.

@@

The second reviewer shared a similar sentiment about not effectively
presenting ideas so we we made improvements to the introduction section to try
and more clearly present our ideas.  We shortened the literature review to only
cite papers that were actually of interest to the paper and tried to clearly
define the two purposes of the paper: to identify inconsistensies with the 2022
reaction times and to decide of 0.1 seconds is a fair barrier. We try to be precise in
our language as we do not have enough evidence to conclude that the sensors were
calibrated differently, or that Seiko (who make the timing devices) is at fault
for the 2022 reaction times being inexplicably faster. Thus we make general
recommendations for World Athletics and note that the reaction times in 2022
were unusually fast with no clear reason.

@@

Thank you for these encouraging words, we have added additional data to both
the GLMM and clusrank section in order to strengthen our arguments: there was
an anomaly in 2022 and that the reaction time barrier of 0.1 seconds is not
grounded in formal statistical analysis.

\jy{State what expansions we
  have done.}
\of{After we are wrapped up with the manuscript it will be easier to identify
all the improvements we made. Let's just leave this here for now.}

@@

We have re-worked the introduction and discussion sections of the paper to more
clearly outline the goals of the paper.  The organization of the paper was
something that both reviewers commented on so we did our best to try and
streamline the two key points of the paper: to decide of 0.1 seconds is fair,
and to identify inconsistensies with the 2022 reaction times. With these two
points emphasized we tried to remove sections that did not support these central
ideas.  In regards to the language surrounding the rank comparison text and the
timing device at the 2022 Championships we are intentionally vague as we do not
have evidence to make a stronger claim.  For example, we cannot claim that the
timing device was faulty as miscalibrated as there is no public information
available about that.  We do not want to overstep and put ourselves or the
publication at risk of defamation claims from Seiko because we called their
devices faulty when they are not.  Instead we suggest to the reader that 2022
was an unusually fast year and comment that there is no explanation as to why
athletes improve reaction times over the span of a few months and let the reader
decide what they think is the underlying cause of the faster reaction times.

@@

We are attempting to answer question 1 and have included reaction times from
additional sprinting events from World Championships.  The inclusion of
additional data was also suggested by the first reviewer and we have provided
our explanation of what data we used there.

\begin{comment}
I am also not convinced that the GLMM methods used for evaluating the
probability of disqualification are valid. The tests use data of reaction times
from performances that were NOT disqualified.  The data is therefore truncated
with no samples from the “left” or “disqualified” tail. This means that any
conclusions about the probability of a reaction time falling below this
threshold is purely extrapolation from the distributional assumptions. The
authors use a gamma distribution based on its fit to the data they have. In my
opinion, the choice of a gamma distribution is not sufficiently justified to
make extrapolating claims about the probability mass in the upsampled tail. I am
not convinced there is enough data to make a strong argument here, especially
when the previous literature on reaction time uses different distributions.
\end{comment}
We will start by addressing the comments related to runners who have been
disqualified.  Some of the runners who were disqualified had a negative reaction
time, indicating they were trying to predict the starting gun and actually left
before the gun even fired.  These runners cannot be included in the analysis
because the point of the GLMM section is to question the current reaction barrier
methods by determining how fast athletes can react in; but athletes with a
negative time left before the gun even went off (Also practically speaking
a negative reaction time makes little sense).  As for the roughly 20 athletes
across men's and women's data without negative times but were disqualified, we
analyzed how the inclusion of these athletes effected our results.  What we found
was that the probability of a reaction time below 0.1 seconds increased from
0.0029 to 0.0039 and we have attached the R code so that you can see our work.
\of{Not sure the technical details of how this is going to work but we will come
back to this before we submit.} Additionally there were similar parameter values
for the standard deviations of the random effects and we decided that we were
going to keep these observations in our data.

As for why the particular model was
chosen, we looked at several types of models and chose the one that we thought
fit the residual qq plot the best and had the largest log likelihood. After
selecting the gamma generalized linear model we again looked at several link
functions such as "identity", "log",

\of{Need to justify why we did what we did and disadvantages of including the
disqualified
Do this in the paper and reply. Say section where we made any changes. Give
justifications for any reasons for disagreeing}

\of{Will clean this up based on Prof Yan and Schifano's comments to 11/5 email}

In terms of the choice of distribution, Brosnan and Harrison used a exponentially
modified gaussian distribution using athletes sex, ruling periods, competition
rounds. The article is now behind a pay wall.
\jy{Not a reason for not citing it. We can still find pdf online.}

\jy{Using log-normal distribution for the reaction times is quite easy
  to do. You just take log of the reaction times, and then model them
  in a GLMM with Guassian family. I think we did this, and decided it
  was not good perhaps from the residual qq plots. Please find them
  and add a discussion into the paper.}

\jy{This references has been cited 31 times. Some of these are closely
  related to our topic and should be cited as appropriate.}

@@

Unfortunately data such as this is not easy to come by.  We found as much as we
could online, but there is simply not enough data to justify a formal analysis
of his reaction times across lower level meets.  Many track and field races do
not record data on reaction times until at the very high level (national
and world championship).  Simply put, reaction times are not heavily researched
and thus the data material in insufficient.
\jy{We have added a discussion in paragraph ??? of the Discussion Section.}

\jy{Every point we cannot answer needs to have some response reflected
  in discussions in the manuscript.}

@@

\begin{comment}I have no idea how to interpret the numbers displayed in Table 2. These
need to be made interpretable, or they add nothing to the paper.
\end{comment}

@@

Yes, this has been resolved.

@@

\subsection*{Associate Editor}
\jy{Put this section before the reviewers.}

We have taken additional steps to streamline the manuscript and enhance the
presentation of our work. Thank you for your time and we appreciate re-considering
our submission.

%\bibliographystyle{chicago}
%\bibliography{citations}
\end{document}

>>>>

\subsection*{To Associate Editor}

Thank you for your feedback. We have carefully addressed the
reviewers' comments and taken additional steps to streamline the
manuscript, focusing on a clearer presentation of the key goals and
findings from the outset. We appreciate your time and consideration in
re-evaluating our submission.

@@

Thank you for the valuable suggestions. We have extended our analysis
by including data from both men's and women's 100-meter dash. This was
justified by a GLMM fit to these data and the event type was not found
to be significantly different in reaction time.

We also considered including additional sprinting events such as the
200-meter dash; however, in a GLMM fit to an expanded dataset that
additionally included 200-meter, the event type effect became
significant. Subsequent pairwise comparisons using Tukey’s Honest
Significant Difference (HSD) test revealed a p-value of 0.0021 when
comparing the reaction times of the 200-meter dash and the 110-meter
hurdles, and a p-value of 0.000017 when comparing the 100-meter dash
and the 200-meter dash. These results indicate that reaction times in
the 200-meter dash are significantly slower, suggesting that this
event does not align well with our focus on extreme reaction times.
Consequently, we decided to exclude the 200-meter data from our
primary analysis.

This outcome is not unexpected, as athletes competing in longer sprint
events, such as the 200-meter dash, may not prioritize reaction time
as highly as those specializing in shorter sprints like the 100-meter
dash. While it is common for amateur and collegiate athletes to
compete across multiple sprint distances, elite athletes at the World
Championship level may not exhibit equally strong performances in all
sprint events. Our study aims to evaluate the probability of observing
extreme reaction times, particularly in the context of the 0.1-second
threshold. Including data from events with systematically slower
reaction times would artificially inflate the extremity of a
0.1-second reaction time, potentially biasing our conclusions.

In the revision, we have added explanation about why 100-meter dash
data for both men and wemen were included, while 200-meter dash data
were not included in our analysis in Section~???, paragraph~???.
The subsequent analyses in the manuscript have been updated with the
expanded dataset that included both men and women's 100-meter dash
data; see updated results in Sections~???.

@@

The heat effect measures the speed-level of the heat.  We added some additional
explanations in the first paragraph of section 3.2 to make it clear what the
venue and heat effect refer to. We did not want to call it a "race effect" as we
were worried that may be interpreted to mean the race of the athlete.  We did
not think it would be informative to  add and effect for individual runner's
average ability as many of observations in the data come from runners
\jy{It's the number of runners with 1 data point that determines how
  hard the athlete effect can be estimated. How many runers had only 1
  data point?}
with very few (1-3) data points, so taking an average of such a small sample may
not be informative, especially if there are outliers or an athlete had a bad
start.

@@

We appreciate the feedback, which was echoed by the second
reviewer. To address these concerns, we revised the introduction to
better structure our arguments. We streamlined the literature review
to focus only on studies directly relevant to our analysis and
clarified the paper’s two main objectives: (1) identifying
inconsistencies in the 2022 reaction times and (2) assessing whether
the 0.1-second threshold is fair. We avoided speculating on sensor
calibration or implicating Seiko due to a lack of direct evidence, and
instead, we provide general recommendations for World Athletics while
noting that the unusually fast reaction times in 2022 remain
unexplained.

@@

Thank you for the positive feedback. We expanded both the rank
comparison and the GLMM sections with additional data to strengthen our
analysis. These enhancements support our findings of an anomaly in
2022 and highlight that the 0.1-second reaction time threshold lacks a
foundation in formal statistical evidence.

@@

We have revised the Introduction and Discussion sections to clearly
outline the goals of the study. Both reviewers noted issues with
organization, so we streamlined the manuscript to emphasize the two
main objectives: (1) identifying inconsistencies in the 2022 reaction
times and (2) assessing whether the 0.1-second threshold is fair. We
removed unrelated sections to keep the focus on these central
points. Regarding the rank comparison and timing devices at the 2022
Championships, we intentionally used cautious language, as there is no
evidence to claim a miscalibration or fault in Seiko’s timing
devices. To avoid potential legal risks, we noted the unusually fast
reaction times in 2022 without making unsubstantiated claims. Instead,
we allow readers to interpret the underlying cause of the anomaly
based on the presented data.

@@

Thank you for the feedback. We are attempting to address question~(1)
and have expanded our analysis to include reaction times from
additional sprint events at the World Championships, including both
men and women's 100-meter data. This broader dataset helps provide a
more comprehensive assessment of whether the venue in 2022 was
unusually fast.

\begin{comment}
I am also not convinced that the GLMM methods used for evaluating the
probability of disqualification are valid. The tests use data of reaction times
from performances that were NOT disqualified.  The data is therefore truncated
with no samples from the “left” or “disqualified” tail. This means that any
conclusions about the probability of a reaction time falling below this
threshold is purely extrapolation from the distributional assumptions. The
authors use a gamma distribution based on its fit to the data they have. In my
opinion, the choice of a gamma distribution is not sufficiently justified to
make extrapolating claims about the probability mass in the upsampled tail. I am
not convinced there is enough data to make a strong argument here, especially
when the previous literature on reaction time uses different distributions.
\end{comment}

Thank you for raising these points. Regarding the exclusion of
disqualified runners, we chose not to include athletes with negative
reaction times, as these indicate an attempt to predict the starting
gun rather than a true reaction. Negative reaction times do not align
with the purpose of the GLMM analysis, which aims to evaluate the
feasibility of the 0.1-second reaction threshold. Including negative
times would introduce results where the reaction occurred before the
gun fired, making them unsuitable for this analysis.

For the approximately 20 disqualified athletes with positive reaction
times, we examined the impact of including them in the analysis. The
probability of observing a reaction time below 0.1 seconds increased
\jy{Are these based on the expanded dataset or the original dataset?}
slightly from 0.0029 to 0.0039, with similar parameter estimates for
the random effects. Given these results, we opted to include these
observations in the dataset.

Distribution: \jy{need to check R package gamlss.}

\jy{Find these references and cite them:
Sprint Start Regulation in Athletics: A Critical Review.
Milloz M, Hayes K, Harrison AJ.
Sports Med. 2021 Jan;51(1):21-31. doi: 10.1007/s40279-020-01350-4.
PMID: 33125639 Review.
Reaction time aspects of elite sprinters in athletic world championships.
Tønnessen E, Haugen T, Shalfawi SA.
J Strength Cond Res. 2013 Apr;27(4):885-92. doi: 10.1519/JSC.0b013e31826520c3.
PMID: 22739331
}

@@

Thank you for this suggestion. Unfortunately, comprehensive data on
Devon Allen's reaction times is not readily available. We gathered as
much data as we could find online, but there is insufficient
information from lower-level meets, as reaction times are typically
only recorded at major competitions (e.g., national and world
championships). Given the limited data, we could not conduct a formal
statistical analysis of his reaction times. However, we have added a
brief discussion of this issue in paragraph~???] of the Discussion Section.

@@

\begin{comment}
I have no idea how to interpret the numbers displayed in Table 2. These
need to be made interpretable, or they add nothing to the paper.
\end{comment}

@@

Yes, this has been rephrased for clarification.

@@

\bibliographystyle{apalike}
\bibliography{citations}
\end{document}
